---
title: "DetailedManuscriptRevisions"
author: "Shannon Rankin"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions

When resubmitting your manuscript, please carefully consider all issues mentioned in the reviewers' comments, outline every change made point by point, and provide suitable rebuttals for any comments not addressed.

To submit your revised manuscript go to <https://www.editorialmanager.com/ecoinf/> and log in as an Author where you will see a menu item called 'Submission Needing Revision'.

Please resubmit your manuscript by Feb 01, 2024.

```{r cars}
summary(cars)
```

## Requested Revisions

-   Research Elements (optional)

This journal encourages you to share research objects - including your raw data, methods, protocols, software, hardware and more -- which support your original research article in a Research Elements journal. Research Elements are open access, multidisciplinary, peer-reviewed journals which make the objects associated with your research more discoverable, trustworthy and promote replicability and reproducibility. As open access journals, there may be an Article Publishing Charge if your paper is accepted for publication. Find out more about the Research Elements journals at <https://www.elsevier.com/authors/tools-and-resources/research-elements-journals?dgcid=ec_em_research_elements_email>.

-   the Introduction does not cover the pertinent literature and fails to highlight the novelty of your study. You are missing important work that has been published on the topic, including the journal Ecological Informatics. This is not a journal that solely focuses on the technical aspects of the published scholarship, we also want to see context, broader implications of your work, critical discussion of the uncertainties etc.

    -   Add a paragraph about different approaches to automated classification, including CNNs, and what the strengths/limitations are. Discuss how even though some of these are really advancing and are becoming increasingly accessible to non-programmers, there is still a high barrier of entry. Banter is stupid easy & can be used by people with limited programming experience. but how does it was developed for working with multiple call types-- how does it work for just clicks?

-   The quality of your Figures is highly problematic and needs to improve significantly.

-   The Discussion is limited and fails to draw parallels with relevant studies from the peer-reviewed literature or put the presented results into the right context. You need to augment significantly the Discussion section in your paper and evaluate your work with reference to similar studies from the field of ecological informatics; several examples are also available in the journal Ecological Informatics.

-   You are missing the Research Highlights, which should not exceed the maximum of 85 characters!!

-   The justification for employing BANTER models is clear. **One could argue that more recent, deep learning-based methodologies could have been incorporated.** However, given that the detection performance in the study appears quite robust, the decision to utilize random forest detectors seems justified, as it doesn't seem to have adversely affected the results. [Discuss value of these methods given advances in CNNs, etc]

-   Improve quality of ALL images

    ## Detailed Responses and Rebuttals

Research Highlights (maybe 3-5 highlights):

-   [Banter is easy to use open source classifier for bioacoustic data to be used on multiple call types]

-   [Banter can be applied to single call types when multiple detectors are used to identify those individual call types]

-   [how well did banter do at classification?]

I would like to thank the reviewers and editors, especially the Associate Editor, for their review of our manuscript.

-   ***Request to include expanded review of research on the research topic (Introduction).*** The associate editor requested that we expand our Introduction to include work that has been published on the topic. Our intention was to focus on a new application of an existing tool, and I assume that the associate editor would like increased consideration of deep learning methods for acoustic classification. To that end, I have included an additional paragraph in the introduction that includes consideration of these methods and a justification of why despite the incredibly advances in deep learning in recent years, consideration of machine learning methods (such as models based on Random Forest) may be preferred for some classification tasks.

-   **Request to evaluate our results and draw parallels with relevant studies in the peer reviewed literature to provide important context (Discussion).** I included a brief comparison of our results to other machine learning methods to acoustically classify beaked whales, and highlighted how the number of classes and the sample size per class differentiates these studies. ***I also included a brief discussion of how the ease of use of BANTER allows biologists to readily adopt these methods, whereas the barrier to entry for most other machine learning models is steep.*** I hesitate to dive deeper into this discussion as these methods are evolving extremely rapidly, and while deep learning methods are rapidly advancing and are arguably the future of our science, at this point they have requirements that are beyond the scope of many scientists. Our focus is on how to use existing tools to provide reasonable results for imperfect data. I have added an additional statement encouraging the preparation of training sets for the eventual consideration in developing deep learning routines.

-   Provide Context, broader implications of our work, and critical discussion of the uncertainties

-   Improve resolution of Figures

-   Include Research Highlights
